---
title: "Homwork_07_11_17"
author: "Lucia Tischer"
date: "9 November 2017"
output: beamer_presentation
fontsize: 8pt
fig_width: 12
fig_height: 10
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, eval = TRUE)
```

## 1) Creating the data

First, I created a data set, consisting of body size, weight and sex of 150 persons:

- 50 men: average height of 180 cm, average weigth of 85 kg (data_men).
- 50 women: average height of 170 cm, average weight of 68 kg (data_women).
- 50 children: average heigth of 140 cm, average weigth of 50 kg (data_children).

```{r, echo = TRUE}
data_men <-  data.frame(sex = "man", 
                        heigth = round(sort(rnorm(50, mean = 180, 
                                                  sd = 10))), 
                        weigth = round(sort(rnorm(50, mean = 85, 
                                                  sd = 10))))
```
```{r, echo = FALSE, eval = TRUE}
data_women <-data.frame(sex = "woman", 
                        heigth = round(sort(rnorm(50, mean = 170, sd = 10))), 
                        weigth = round(sort(rnorm(50, mean = 68, sd = 10))))
data_children <-data.frame(sex = "child", 
                          heigth = round(sort(rnorm(50, mean = 140, sd = 10))), 
                          weigth = round(sort(rnorm(50, mean = 50, sd = 10))))
```


I connected the data to one data-frame:
```{r, echo = TRUE}
names(data_men) <- colnames(data_men)
names(data_women) <- colnames(data_men)
names(data_children) <- colnames(data_men)
data_complet = rbind(data_men, data_women, data_children)
```

***
To check the data each 30th entry, I define a squence from 30 to 150 including 1 and query the data at these points:
```{r, echo = TRUE}
position = c(1, seq(30, length(data_complet$heigth), 30))
data_complet[(position), ]
```


## 2) Saving the data as csv-file and reloading the data
I saved the data.frame as csv-file...
```{r, echo =TRUE, eval = FALSE}
write.table(data_complet, file ="~/data_homework_07_11.csv", 
            sep = ";", 
            row.names = FALSE)
```

```{r, echo = FALSE, eval = TRUE, message = FALSE}
write.table(data_complet, file ="C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/R-Skripte/geostat_homework/data_homework_07_11.csv", sep = ";", row.names = FALSE)
```

... and reloaded it.
```{r, echo =TRUE, eval = FALSE}
table = read.csv("~/data_homework_07_11.csv", 
                 sep = ";", 
                 header = TRUE)
```

```{r, echo = FALSE, eval = TRUE, message = FALSE}
table = read.csv("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/R-Skripte/geostat_homework/data_homework_07_11.csv", 
                 sep = ";", 
                 header = TRUE)
```

## 3) Queriing data points and information
Getting some information about the data, for example:
```{r, echo = TRUE, results = FALSE}
summary(table) # overview over all values.
summary(table[table$sex == "man", ]) # ... over values of men.
str(table) # summary of included variables and their classes.
mode(table) # storage mode of "table" --> "list"
head(table) # first 6 lines of "table".
# query single values, like: 
# how many weight-data points of children do exist.
length(table$weigth[table$sex == "child"])
# maximum heigth in children.
max(table$heigth[table$sex == "child"]) 
# how many people between 160 and 180 cm.
length(table$heigth[table$heigth >= 160 & table$heigth <= 180]) 
```

## 4) Plotting the data
Compare the data points visually by plotting:  

- Scatterplot:
```{r}
par(mfrow =c(1, 1))
plot(table$heigth[table$sex == "man"], table$weigth[table$sex == "man"],
     col = "blue", 
     xlab = "body size [cm]", 
     ylab = "body weigth [kg]", 
     cex.lab = 1.3, 
     las = 1, 
     pch = 20, 
     xlim = c(95, 210), 
     ylim = c(25, 110))
points(table$heigth[table$sex == "woman"], table$weigth[table$sex == "woman"],
     col = "darkred", 
     pch = 20)
points(table$heigth[table$sex == "child"], table$weigth[table$sex == "child"],
     col = "orange", 
     pch = 20)
legend(x = 95, y = 110, 
       legend = c("men", "women", "children"), 
       col = c("blue", "darkred", "orange"), 
       fill = c("blue", "darkred", "orange"))
```

***

-  Barplot
```{r}

boxplot(table$weigth ~ table$sex, 
        names = c("children", "men", "women"), 
        xlab = "persons", 
        ylab = "body weigth [kg]", 
        cex.lab = 1.3, 
        las = 1)
```

## 5) "Cutting" parts of the data to check
How many women are in each body-heigth group?
How is the distribution pattern?

- Check minimum and maximum sizes.
```{r}
max_size_woman <- max(table$heigth[table$sex =="woman"])
min_size_woman <- min(table$heigth[table$sex =="woman"])
```

- Group the data:
```{r,  echo = TRUE, eval = TRUE, message = FALSE}
women_sizegroups <- cut(table$heigth[table$sex == "woman"], 
                        breaks = c(150, 160, 170, 180, 190, 200), 
                        include.lowest = FALSE,
                        rigth = TRUE, 
                        labels = c("150 to 160", "161 to 170", 
                                   "171 to 180", "181 to 190", 
                                   "above 190"))
table(women_sizegroups)
```

***

- Plot the number of women per group
```{r, echo = FALSE, eval = TRUE, message = FALSE}
plot(women_sizegroups, 
     ylim = c(0, 25), 
     xlab = "body sizes women [cm]", 
     ylab = "number of women per group", 
     cex.lab = 1.3, 
     las = 1)
```

***

- show the distribution amongst the size groups,  
  "50 % of the women are smaller than ... cm ..."
```{r, echo = FALSE, eval = TRUE}
women_size_dist <- (quantile((table$heigth[table$sex == "woman"]), 
                                      c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)))
test <- (quantile((table$heigth[table$sex == "woman"]), 
                                      c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)))
plot(women_size_dist, 
     cex.lab = 1.3, 
     xaxt = "n", 
     xlab = "`percent of smallest women´", 
     ylab = "body size [cm]", 
     las = 1)
axis(side = 1, at=c(1,2,3,4,5,6,7,8, 9), 
     labels=c(10, 20, 30, 40, 50, 60, 70, 80, 90),
     las=1)
segments(x0 = 0, y0 = women_size_dist[1], x1 = 1, y1 = women_size_dist[1], col="red")
segments(x0 = 0, y0 = women_size_dist[2], x1 = 2, y1 = women_size_dist[2], col="red")
segments(x0 = 0, y0 = women_size_dist[3], x1 = 3, y1 = women_size_dist[3], col="red")
segments(x0 = 0, y0 = women_size_dist[4], x1 = 4, y1 = women_size_dist[4], col="red")
segments(x0 = 0, y0 = women_size_dist[5], x1 = 5, y1 = women_size_dist[5], col="red")
segments(x0 = 0, y0 = women_size_dist[6], x1 = 6, y1 = women_size_dist[6], col="red")
segments(x0 = 0, y0 = women_size_dist[7], x1 = 7, y1 = women_size_dist[7], col="red")
segments(x0 = 0, y0 = women_size_dist[8], x1 = 8, y1 = women_size_dist[8], col="red")
segments(x0 = 0, y0 = women_size_dist[9], x1 = 9, y1 = women_size_dist[9], col="red")
```

## 6) The functions order(), rank() and sort()

First, I defined an (unordered) example vector of values and tested the functions.
```{r, echo = TRUE,  results = TRUE}
test <- c(2,5,4,3,1,0)
order(test) 
rank(test)  
sort(test)   
```

- order(): shows the order, in which values needed to be  
           to be ascending
- rank(): shows "ranks" of the values  
          (here: first number is 3th highest,  
                 second number is 6th highest, ...)
- sort(): sorts the values in ascending order

