---
title: "Bivand et al. 2013 - Chapter 2"
author: "Lucia Tischer"
date: "28 November 2017"
output: beamer_presentation
fontsize: 8pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Classes and Methods in R

- R as simple calculator: therefore used mathmatical functions/operators are already defined!
- Operators can be used to define sequences.
- Results can be printed.
- Variable contain objects of particular classes.  
  --> the class determines the method to handle the object
- These objects are stored in particular modes.

***
Example 1:

```{r}
# define a variable using predefined operators
# (here: integer sequence)
x <- pi * (1:10)^2  
```

```{r}
x         # print the variable (by default)
print(x)  # print the variable (explicit method)

class(x)  # the object is of class numeric
typeof(x) # storing mode is double
```


***

If no method is defined for a class,  
it´s passed to a default method = generic function.

Example 2:
```{r}
class(cars) # the object is of class data.frame
typeof(cars) # it contains values in a list (vector of arbitrary objects)
```

***

The methods names() and summary() can be applied:
```{r}
names(cars) # cars contains values from two variables
summary(cars) # get information about the data structure
str(cars) # get summed up information, including size and class of the object
```


***

Dataframes view data in rectangles:
- rows of observations
- columns of variables

Rownames

## Spatial Objects

```{r cars, echo = TRUE}
summary(cars)
```

## Spatial points

```{r pressure}
plot(pressure)
```

## Spatial Lines


## Spatial Polygons


## Spatial Grid and Spatial Pixel Objects


## Raster Objects and the raster package
