---
title: "Bivand et al. 2013 - Chapter 2"
author: "Lucia Tischer"
date: "28 November 2017"
output: beamer_presentation
fontsize: 8pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/R-Skripte/geostat_homework")
```

## Introduction



## Classes and Methods in R

- R as simple calculator: therefore used mathmatical functions/operators are already defined!  


- Operators can be used to define sequences.  


- Results can be printed.  


- Variable contain objects of particular classes.  
  --> the class determines the method to handle the object  
  
  
- These objects are stored in particular modes.

***

Example:

```{r}
# define a variable using predefined operators
# (here: integer sequence)
x <- pi * (1:10)^2  
```

```{r}
x         # print the variable (by default)
print(x)  # print the variable (explicit method)

class(x)  # the object is of class numeric
typeof(x) # storing mode is double
```


***

If no method is defined for a class,  it´s passed to a default method.  

= generic function. 


Example:
```{r}
class(cars) # the object is of class data.frame
typeof(cars) # it contains values in a list (vector of arbitrary objects)
```

***

The methods names() and summary() can be applied on *cars*:
```{r}
names(cars) # cars contains values from two variables
summary(cars) # get information about the data structure
str(cars) # get summed up information, including size and class of the object
```


***
--> *cars* is of class *data.frame*  

Dataframes view data in rectangles:  

- rows of observations
- columns of variables

Adressing rownames:
```{r}
row.names(cars) # query the rownames
test_frame <- cbind(c(1:10), c(rnorm(10)))
row.names(test_frame)
row.names(test_frame) <- c(1:10) # assign rownames
row.names(test_frame)
```

***

*formula* = abstraction to facilitate statistical modelling  

- 2-sided, response- and determining variable
- used for model fitting functions
- accompanied by *data* argument: where to find information
```{r}
class(cars$dist ~ cars$speed)
lm(dist ~ speed, data = cars)
```

***
  
A *formula* can be used for factors as well.
```{r}
# create a factor for speed by cutting into quartiles
cars$qspeed <- cut(cars$speed, 
                   breaks = quantile(cars$speed),
                   include.lowest = TRUE)
is.factor(cars$qspeed)
```

--> the plot method will choose the right method depending on the given class

***
```{r, echo = FALSE, eval = TRUE, message = FALSE}
par(mfrow = c(1,2))
plot(dist ~ speed, data = cars, 
     main = "numeric: scatterplot", 
     cex.main = 1.5)
plot(dist ~ qspeed, data = cars, 
     main = "factor: boxplot", 
     cex.main = 1.5)
```

--> the plot method will choose the right method depending on the given class

***
How does the *lm* funcion react to the factor?
```{r}
lm(cars$dist ~ cars$qspeed)
```

--> a dummy variable form is created automatically.


## Spatial Objects
```{r warning = FALSE, echo = FALSE}
library(sp)
```

```{r}
getClass("Spatial")
```

***

--> Spatial Objects contains two slots:  

- a matrix with 2 rows and columns for the coordinates  
- a CRS class object, defining the coordinate reference system  

Spatial subclasses are also defined by the class *Spatial*.
  
***

```{r}
getClass("CRS")
```

--> only slot value: character string
simpliest string for geographical coordinates: "+proj=longlat"

***
--> a simple *Spatial* object can be built from a bounding box matrix and a missing coordinate reference system:
```{r}
m <- matrix(c(0, 0, 1, 1), 
            ncol = 2, 
            dimnames = list(NULL,c("min", "max"))) # matrix is defined
crs <- CRS(projargs = as.character(NA)) # CRS is defined as `missing´
crs
# Spatial object is created from the matrix and the CRS
S <- Spatial(bbox = m, proj4string = crs) 
S
```


## Spatial points

- a 2D pint can be discribed by a pair of numbers (x, y)
- geographic coordinates define a unique point, if the ellipsoid model and the prime meridian are also known.  

Check the Class **SpatialPoints**
```{r}
getClass("SpatialPoints")
```
--> A third slot *coords* is added  
--> point coordinates can be added.

***
```{r, eval = TRUE, echo = FALSE, message = FALSE}
CRAN_df <- read.table("CRAN051001a.txt", header = TRUE)
CRAN_mat <- cbind(CRAN_df$long, CRAN_df$lat)
row.names(CRAN_mat) <- 1:nrow(CRAN_mat)
llCRS <- CRS("+proj=longlat +ellps=WGS84")
CRAN_sp <- SpatialPoints(CRAN_mat, proj4string = llCRS)
```

*SpatialPoints* objects have a summary method:
```{r}
summary(CRAN_sp)
```
--> Object is not projected
--> there are 54 rows of coordinates

***
### Methods for SpatialPoints

Access values of the slots: **bbox** method
```{r}
bbox(CRAN_sp) 
```
--> bounding box of the object is returned.
--> first row: west-east, second row: south-north direction.

Reporting to current projection string: **proj4string** method
```{r}
proj4string(CRAN_sp)
proj4string(CRAN_sp) <- CRS(as.character(NA)) # set to NA
proj4string(CRAN_sp) <- llCRS # re-set to old value
```

***

Extracting the coordinates: **coordinates** method
```{r, eval = TRUE, echo = TRUE, results=FALSE}
brazil <- which(CRAN_df$loc == "Brazil") # choose the locations in Brazil
brazil
```

```{r}
# choose coordinates coresponding to locations
coordinates(CRAN_sp)[brazil, ] 
```

***

Access the object by indexing:
```{r}
summary(CRAN_sp[brazil, ])
```

--> object ist of class *SpatialPoints*
--> new bounding box
--> projectin is retained

***
### Data Frames for Spatial Point Data
--> data from different sources can be mereged, if matrix rownames (coordinates) and data frame rownames (measurements) match  

--> data frame get re-ordered to suit coordinate points  

Create a **SpatialPointsDataFrame** from a **SpatialPoints** object and a **data frame**
```{r}
CRAN_spdf1 <- SpatialPointsDataFrame(CRAN_mat, 
                                     CRAN_df, 
                                     proj4string = llCRS, 
                                     match.ID = TRUE)
CRAN_spdf1[54, ]
```
--> the two objects got merged by their rownames.

***
--> same would result, if the data frame got mixed randomly.
```{r}
s <- sample(nrow(CRAN_df))
CRAN_spdf2 <- SpatialPointsDataFrame(CRAN_mat, 
                                     CRAN_df[s,], 
                                     proj4string = llCRS, 
                                     match.ID = TRUE)
all.equal(CRAN_spdf2, CRAN_spdf1)
```
--> ID values need to match, otherwise: ERROR.

***

Examine the contents of a **SpatialPointsDataFrame**
```{r}
getClass("SpatialPointsDataFrame")
```
--> slot *data* keeps the information (in a **data.frame** object).  
--> also contains information contained in *Spatial* class object like bbox or proj4string.  
--> Behavior of **SpatialDataFrames** +/- like data frames: methods like *names()* or *str()* are included.


***
Other possibility to construct a **SpatialPointsDataFrame**:
```{r}
# a SpatialPoints object as first argument in the function
CRAN_spdf4 <- SpatialPointsDataFrame(CRAN_sp,
                                     CRAN_df)


all.equal(CRAN_spdf4, CRAN_spdf2) # check if the result is equal.
```
***

Assign coordinates to a data frame:  

--> original data frame is modified!  
  
  
```{r}
CRAN_df0 <- CRAN_df # dublicate the data frame

# assign coordinates by matrix of coordinates
coordinates(CRAN_df0) <- CRAN_mat

proj4string(CRAN_df0) <- llCRS # define projection of coordinates

all.equal(CRAN_df0, CRAN_spdf2) # check if objects are equal
```

***
**SpatialPoints**: can also represent lines, if information is associated with point, rather than the line.  
--> sequence numbers for the points need to be entered  
--> tracing in order is possible

```{r}
# read in the data
turtle_df <- read.csv("seamap105_mod.csv", sep = ";")

# observtions get timstamped
timestamp <- as.POSIXlt(strptime(as.character(turtle_df$obs_date),
                                 "%m/%d/%Y%H:%M:%S"), "GMT")
# new data frame with data and timestamp
turtle_df1 <- data.frame(turtle_df, timestamp = timestamp)

# longitude values are corrected to positiv ones (?!)
turtle_df1$lon <- ifelse(turtle_df1$lon < 0, 
                         turtle_df1$lon ++ 360, 
                         turtle_df1$lon)

# new data frame is re-ordered by the timestamp
turtle_sp <- turtle_df1[order(turtle_df1$timestamp), ]

# coordinates and crs are defined
coordinates(turtle_sp) <- c("lon", "lat")
proj4string(turtle_sp) <- CRS("+proj=longlat +ellps=WGS84")
```


## Spatial Lines
--> lines as sequences/set of points, skipping the NA-values.

Compare **Line**, **Lines**, and **SpatialLines** objects:
```{r}
getClass("Line")
```
--> **Line** object is a matrix of 2D coordinates  

--> doesn´t inherit from *Spatial* class  

***

```{r}
getClass("Lines")
```
--> **Lines** object consists of list of **Line** objects  

--> doesn´t inherit from *Spatial* class  

***

```{r}
getClass("SpatialLines")
```
--> inherits from **Spatial** class  
(bbox- and proj4string slots contain spatial information for lines slot)

***

Example for a **SpatialLines** object:
```{r, message = FALSE}
library(maps) # load the package (containing the data)
japan <- map("world", "japan", plot = FALSE) # define the example data
p4s <- CRS("+proj=longlat +ellps=WGS84") # define the CRS
library(maptools) # load the package (containing the function)
SLjapan <- map2SpatialLines(japan, proj4string = p4s) # convert the data
str(SLjapan, max.level = 2) # examine SpatialLines object
```

***

**SpatialLines** objects are similar to **SpatilaPolygons** objects.  

--> contents of objects can be explored with *lapply()* or *sapply()*  

--> What´s the length of the **Lines** slot of each object in the **Lines** slot of SLJapan?
```{r}
Lines_len <- sapply(slot(SLjapan, "lines"), 
                    function(x) length(slot(x,"Lines")))
table(Lines_len)
```
--> None of the 34 **Lines** slots contains more than 1 **Line** object.

***
Other possiblity to create **SpatialLinesDataFrame**:  
*ContourLines2SLDF()* function  

--> *slot()* can be used to look inside the slots of objects.
```{r}
volcano_sl <- ContourLines2SLDF(contourLines(volcano))
t(slot(volcano_sl, "data"))
```

***
Reading in data and format it to a **SpatialLines** object directly:
```{r}
llCRS <- CRS("+proj=longlat +ellps=WGS84")
auck_shore <- MapGen2SL("auckland_mapgen.dat.txt", llCRS)
summary(auck_shore)
```


## Spatial Polygons

Polygon = closed line.  
--> first point of sequence = last point of the sequence.  
--> Multiple polygons are separated by NA values.  

Which of the imported lines represent shorelines of islands?
```{r, eval = TRUE, echo = TRUE, message = FALSE}
lns <- slot(auck_shore, "lines")
table(sapply(lns, function(x) length(slot(x, "Lines"))))
```
--> the slots each contain only one object.  


***
  
How many of the Lines are Polygons?
```{r}
islands_auck <- sapply(lns, 
                       function(x) {crds <- slot(slot(x, "Lines")[[1]], "coords")
# are first and last coordinate identical?                       
identical(crds[1, ], crds[nrow(crds), ])})
table(islands_auck)
```
--> 64 objects have identical first/last coordinates  
--> are polygons!

***
  
Compare classes **Polygon** and **Line**:
```{r}
getClass("Polygon")
```
--> **Polgon** extends **Line**  

- labelpoint: centroid of the polygon  
- area of the polygon  
- wether polygon is declared as a hole or not  
- ring direction of the polygon  
  
  --> polygons are **LinearRing** objects (closed LineString)

***

How about the **Polygons** class?
```{r}
getClass("Polygons")
```
--> **Polygons** objects are *MultiPolygon* objects  
--> contain a list of **Polygon** objects  
--> additional plot order slot

***

Finally: **SpatialPolygons** objects:
```{r}
getClass("SpatialPolygons")
```
--> **SpatialPolygons** are sets of **Polygons** objects with additional slots of **Spatial** objects  
(bounding box, projection information of the set)  
--> also: additional plot order slot  
(plot order decreases from polygon with biggest to smallest area)

Creation of **SpatialPolygons** objects: from shoreline data with closed lines

***

### 2.6.1 SpatialPolygonsDataFrame Objects

--> spatial representation of polygons **+** data  

--> **SpatialPolygons** objects have tags 

--> these tags are matched with row names of the data frame  
(one-to-one matching)

--> correct data rows are associated with the correct spatial objects  

If tags and row names don´t correspond: **ERROR**  
--> **SpatialPolygonsDataFrame** objects can´t be created.

***  

Example: associate test results with country boundaries:  

```{r}
library(maps)

# define boundaries
state.map <- map("state", plot = FALSE, fill = TRUE)
IDs <- sapply(strsplit(state.map$names, ":"), 
              function(x) x[1])

library(maptools)

# convert boundaries to a SpatialPolygons object
state.sp <- map2SpatialPolygons(state.map,
                                IDs = IDs,
                                proj4string = CRS("+proj=longlat +ellps=WGS84"))
```
***

The data frame rows are suited to the **SpatialPolygons** by identifying tag matching:
```{r, result = FALSE}
# read in the test data
sat <- read.table("state.sat.data_mod.txt", 
                  row.names = 5,
                  header = TRUE)
# match tags and row names
id <- match(row.names(sat), row.names(state.sp))

sat1 <- sat[!is.na(id), ]
state.spdf <- SpatialPolygonsDataFrame(state.sp, sat1)
```

***
Have a look on the **SpatialPolygonsDataFrame**:
```{r}
str(state.spdf, max.level = 2)
```
***

### 2.6.2 Holes and Ring Direction (?)

- Polygons can be marked as holes (f.e. TRUE/FALSE: hole slote)

- hole slot can be set directly while reading in the data:  
--> four levels: 1-land, 2-lake, 3-island in lake, 4-lake on island in lake


## 2.7 Spatial Grid and Spatial Pixel Objects (?)
```{r}
getClass("GridTopology")
```
--> Grid = regular objects  
--> much less information is required to define their structure  
--> extent is fiven by cell resolution and numbers of rows and columns



- - - example data not found ...

## 2.8 Raster Objects and the raster package
```{r, message =FALSE}
library(raster)
r <- raster("70042108.tif") # define a raster object
class(r) # query some information about the object
inMemory(r)
object.size(r)
cellStats(r, max)
cellStats(r, min)
```
*** 

How to remove values of less then or equal to zero:
```{r, message=FALSE, echo = TRUE, eval =TRUE}
out <- raster(r) # new file gets defined with same structure as r
bs <- blockSize(out) # define number of entire row/chunk sizes

# open new file and start writing the values
out <- writeStart(out, filename = tempfile(), overwrite = TRUE)
for (i in 1:bs$n) { 
  v <- getValues(r, row = bs$row[i], nrows = bs$nrows[i]) 
  v[v <= 0] <- NA
  writeValues(out, v, bs$row[i])
}
out <- writeStop(out) # close file/stop writing
```
***
Check the new minimum and maximum values:
```{r}
cellStats(out, min)
cellStats(out, max)
```

***
Possibility for plotting raster data:
```{r}
plot(out, col = terrain.colors(100))
```
***

Coercion between classes of the **sp** and **raster** package is possible.  
--> Coercion from raster layer to SpatialGridDataFrame:
```{r}
r1 <- as(out, "SpatialGridDataFrame")
summary(r1)
```
***

--> Coercion from SpatialGridDataFrame to raster layer:
```{r}
r1 <- as(out, "SpatialGridDataFrame")
summary(r1)
```


